Aligning marine species range data to better serve science and conservation

### 2016-01-31 data_explore/data_prep.Rmd
tracking down the discrepancies in species numbers

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
                           stringsAsFactors = FALSE) %>%
  filter(!str_detect(spatial_source, 'subpop')) %>%
  filter(!str_detect(spatial_source,  'alias')) %>%
  select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
         popn_category, spp_group, id_no, spatial_source)

write.csv(spp_all_remote, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)

### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
### that also have a non-NA AquaMaps ID code.
spp_am_v_iucn <- spp_all_remote %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  unique()
write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
          row.names = FALSE)

### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_marine <- spp_iucn_marine %>%
  select(sciname, iucn_sid, iucn_category = category, popn_trend, parent_sid, subpop_sid)
write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
          row.names = FALSE)

### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_maps <- spp_iucn_maps %>%
  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
  filter(id_no %in% spp_iucn_marine$iucn_sid)   ### ditch terrestrial and freshwater spp
write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
          row.names = FALSE)

>>>>

```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
                           stringsAsFactors = FALSE) 
spp_all_remote_clean <- spp_all_remote %>%
  filter(!str_detect(spatial_source, 'subpop')) %>%
  filter(!str_detect(spatial_source,  'alias'))
### At this point, plenty of duplicates:
### * This code returns 518 duplicated names:
###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
### * some are AM subpops which were ignored in processing the remote spp_all file.
###   Note: all subpops have an identified parent, so can filter them out without
###   losing any species.
###   code to verify this:
###     z_parent <- spp_all_remote_clean %>% 
###       filter(!is.na(subpop_sid) & spatial_source == 'am')
###     z_subpop <- spp_all_remote_clean %>% 
###       filter(!is.na(parent_sid) & spatial_source == 'am')
###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
###   eliminated when select(-iucn_subpop) column and then unique()

### filter out spatial_source == 'am' that also have parent_sid (i.e.
### these are AM-mapped subpopulations); then: 
### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
### these are creating false unique rows.
spp_all <- spp_all_remote_clean %>%
  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
         popn_category, spp_group, id_no, spatial_source) %>%
  unique()
### At this point, still some duped scinames - these are ones with multiple IUCN
### SIDs that are not designated as parents or subpops.  We will brute-force
### this by just eliminating the duplicated rows, keeping only the first instance.
### By arranging by iucn_sid, we will keep the lowest value, which seems to
### generally correspond to the parent...
spp_all <- spp_all %>%
  arrange(iucn_sid)
spp_all <- spp_all[!duplicated(spp_all$sciname), ]

length(spp_all$sciname) - length(unique(spp_all_remote_clean$sciname))
### no duped scinames - same length as original (cleaned) list

write.csv(spp_all, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)

```{r get_other_vars, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
### that also have a non-NA AquaMaps ID code.
spp_all <- read.csv(file.path(dir_data, 'spp_all.csv'), stringsAsFactors = FALSE)

spp_am_v_iucn <- spp_all %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  unique()
write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
          row.names = FALSE)

### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_marine <- spp_iucn_marine %>%
  dplyr::select(sciname, iucn_sid, iucn_category = category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
          row.names = FALSE)

### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
                            stringsAsFactors = FALSE) %>%
  rename(iucn_sid = id_no)

spp_iucn_maps <- spp_iucn_maps %>%
  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
  filter(iucn_sid %in% spp_iucn_marine$iucn_sid) %>% ### ditch terrestrial and freshwater spp
  unique() %>%
  arrange(iucn_sid)
spp_iucn_maps <- spp_iucn_maps[!duplicated(spp_iucn_maps$sciname), ]
write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
          row.names = FALSE)

### 2016-01-31 manuscript-main.Rmd
tracking down the discrepancies in species numbers

@@

data_file <- 'data/spp_list_w_area_trimmed.csv'
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)

>>>>

if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')

@@

data_file <- 'data/spp_list_w_area_trimmed.csv'
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
#  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
